Estimating Neighborhood Rents with Scraped Data

### 2018-06-03 manuscript/index.Rmd
progress on site with multi-page

---
title: "Estimating Neighborhood Rents with Scraped Web Data"
author: "Chris Hess"
date: "June 2, 2018"
output: html_document
abstract: "Online classified ads are a key source of information about local rental housing markets. This research asks what temporal and spatial dynamics explain neighborhood variations in one-bedroom rents in Seattle, WA, and contributes a novel approach to forecasting quarterly neighborhood rent levels using Bayesian models estimated with integrated nested Laplace approximations (INLA). All models were trained on data for 2017 Q2 through 2018 Q1, with forecasts for 2018 Q2 compared against the observed-to-date data for the current quarter (i.e. 2018 Q2). The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
---

@@

#load ACS data extract (codebook in folder)
census <- read_csv(file = "./input/acsExtract.csv")
censusRent <- read_csv(file = "./input/acsRent.csv")
census <- inner_join(census, censusRent)

#read in tract shapefiles for seattle
sea_shp <- readOGR(dsn = "./input/sea_tract_2010/sea_tract_2010.shp",
                  layer = "sea_tract_2010",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F)

#go ahead and save separate versions for graphics later
tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
                  layer = "sea_tract_2000",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F) 
tract2010 <- sea_shp

#source data_load()
source("./helpers/load_data.R")

@@

```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/map_point_density.R") #if config, refresh plot pdf
}
```

![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](./output/map_point_density.png)
<br>

@@

```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/rent_lineplot.R") #if config, refresh plot pdf
}
```

![Figure 2. Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018. Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e. respective 1-Year Estimates for each year period).](./output/rent_lineplot.png)

```{r rent_map_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/rent_map_across.R") #if config, refresh plot pdf
}
```

![Figure 3. Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](./output/rent_map_across_1B.png)

```{r rent_scatter_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../data/cl/craigslistDB.sqlite")){
  source("./helpers/rent_scatter_across.R") #if config, refresh plot pdf
}
```

![Figure 4. Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](./output/rent_scatter_across_1B.png)
<br>

>>>>

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess"
date: "June 2, 2018"
output: html_document
navbar:
  - {title: "hesscl.com", href: "http://hesscl.com", align: right}
  - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: left}
abstract: "Online classified ads are a key source of information about local rental housing markets. This research asks what temporal and spatial dynamics explain neighborhood variations in one-bedroom rents in Seattle, WA, and contributes a novel approach to forecasting quarterly neighborhood rent levels using Bayesian models estimated with integrated nested Laplace approximations (INLA). All models were trained on data for 2017 Q2 through 2018 Q1, with forecasts for 2018 Q2 compared against the observed-to-date data for the current quarter (i.e. 2018 Q2). The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."

---

@@

#load ACS data extract (codebook in folder)
census <- read_csv(file = "../input/acsExtract.csv")
censusRent <- read_csv(file = "../input/acsRent.csv")
census <- inner_join(census, censusRent)

#read in tract shapefiles for seattle
sea_shp <- readOGR(dsn = "../input/sea_tract_2010/sea_tract_2010.shp",
                  layer = "sea_tract_2010",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F)

#go ahead and save separate versions for graphics later
tract2000 <- readOGR(dsn = "../input/sea_tract_2000/sea_tract_2000.shp",
                  layer = "sea_tract_2000",
                  GDAL1_integer64_policy = TRUE,
                  stringsAsFactors = F) 
tract2010 <- sea_shp

#source data_load()
source("../helpers/load_data.R")

@@

```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/map_point_density.R") #if config, refresh plot pdf
}
```

![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
<br>

@@

```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/rent_lineplot.R") #if config, refresh plot pdf
}
```

![Figure 2. Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018. Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e. respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)

```{r rent_map_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/rent_map_across.R") #if config, refresh plot pdf
}
```

![Figure 3. Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)

```{r rent_scatter_across, echo=FALSE, message=FALSE, warning=FALSE}
#determine config (dev or github)
if(file.exists("../../data/cl/craigslistDB.sqlite")){
  source("../helpers/rent_scatter_across.R") #if config, refresh plot pdf
}
```

![Figure 4. Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)
<br>
